---
title: "Longevity of Olympic Athletes"
format: html
editor: visual
execute:
  echo: false
---

```{r}
#| message: FALSE
#| label: setup
library(tidyverse)
library(bayesplot)
library(tidybayes)
library(brms)
```

```{r}
#|message: FALSE
#|label: dataset
ds <- read_csv("athlete_events.csv", show_col_types = FALSE)

```

```{r}
#|label: athlete data

athlete_data <- ds |>
  mutate(medal = if_else(is.na(Medal), 0, 1)) |>
  group_by(ID, Sport, Team) |>
  summarise(
    Total_Medals = sum(medal, na.rm = TRUE),
    Olympics_Attended = n_distinct(Games),
    .groups = "drop"
  )


```

```{r}
#| cache: true
# Define the model formula
formula <- bf(Olympics_Attended ~ Total_Medals + (1|Sport) + (1|Team))

smaller_athlete_data <- slice_sample(athlete_data, prop = 0.01)
# Fit the model
```

```{r}
#|cache: TRUE
#|message: FALSE
#| warning: FALSE
#| results: 'hide'
 fit <-  brm(
  formula = formula,
  data = smaller_athlete_data,
  family = poisson(), 
  seed = 9, 
  refresh = 0, 
  silent = 2)
```

```{r}
#| cache: TRUE
# Check model fit and convergence
#summary(fit)
#plot(fit)
#posterior_predict(fit)
#fixef(fit)

teams <- unique(smaller_athlete_data$Team)
sports <- unique(smaller_athlete_data$Sport)
total_medals <- seq(0, max(smaller_athlete_data$Total_Medals)) # Adjust the range as needed

# Create a tibble with all combinations
all_combinations <- expand_grid(
  Team = teams,
  Sport = sports,
  Total_Medals = total_medals,
  Olympics_Attended = 1:10
)
# Sample a subset of the combinations
sampled_combinations <- all_combinations |> sample_n(1000) # Adjust the number as needed

# Generate predictions using the comprehensive tibble
predictions <- fit |> add_epred_draws(newdata = sampled_combinations, seed = 9)
swimming_predictions <- predictions |> filter(Sport == "Swimming")
athletics_predictions <- predictions |> filter(Sport == "Athletics")
team_ball_sports_predictions <- predictions |> filter(Sport == "Basketball" | Sport == "Volleybal" | Sport == "Handball" | Sport == "Football" | Sport == "Water Polo")

combined_predictions <- bind_rows(
  swimming_predictions %>% mutate(Sport = "Swimming"),
  athletics_predictions %>% mutate(Sport = "Athletics"),
  team_ball_sports_predictions %>% mutate(Sport = "Team Ball Sports")
)
```

```{r}
#|message: FALSE
ggplot(combined_predictions, aes(x = Total_Medals, y = Olympics_Attended, color = Sport)) +
  geom_point() +  # Adjust point size and transparency
  facet_wrap(~ Sport) +
  labs(
    title = "Predicted Number of Olympics Attended Based on Medals Earned",
    x = "Total Medals",
    y = "Predicted Number of Olympics Attended"
  ) +
  scale_color_manual(values = c("Swimming" = "green", "Athletics" = "red", "Team Ball Sports" = "purple")) +
  theme_minimal() +
  theme(
    legend.title = element_blank(),  # Remove legend title
    legend.position = "bottom"       # Position legend at the bottom
  ) + scale_y_continuous(breaks = c(1:10, 2))
```

```{r}
#|message: FALSE
# In first .qmd file 
saveRDS(fit, file = "fit_model.rds")

```

This project explores the relationship between a number of medals and Olympic athlete has earned, and the number of Olympics they have attended. Attempting to explain the longevity of an athlete's Olympic career, we show over the past 120 years, the average medal per Olympics attended ratio, with swimming as an outlier, since it is easy to rack up the medals in swimming. Using Sport and Team as random variable we use a formula of Total_Medals ~ Olympics_Attended + Sport + Team.

